Kinematics of the Palomar 5 stellar stream from RR Lyrae stars

### 2019-08-22 paper/main.tex
Updates from Overleaf

RR Lyrae stars (RRLs) are pulsating horizontal branch giants intrinsically more luminous than main sequence turn-off stars. Although about $\sim 10,000$ time less numerous than main sequence stars, RRLs are extremely useful probes of the Galactic halo as they are well established as standard candles; they can be observed to large distances \citep{Medina2018,Sesar2017c}; and are reliably identified based on their photometric variability, with little to no contamination from other types of variable stars \citep[e.g.]{Holl2018,Drake2017,Mateu2012}.  Their distances can be estimated based on photometric data alone, with errors that go from $\sim8\%$ in the optical to as low as 3\% in the infrared \citep{Neeley:2017, Sesar:2017a}.

Previous studies have found RRLs in Pal5's tails: the first two were found by \citet{Vivas2005} in the QUEST survey; and more recently \citep{Ibata:2017} found about ten new RRLs using the catalogue of PS1 candidate RRLs from \citet{Hernitschek2015}. At its current distance, Pal~5's horizontal branch is bright enough ($\G\sim17.3$) to be observed by \Gaia. Therefore, combining \Gaia~DR2 proper motions with the newer PS1 RRL catalogue from \citet{Sesar2017b} --with improved classification probabilities and light curve periods--  will allows to make precise distance and proper motion measurements along the full extent of the Pal~5 stream. Both, key ingredients for dynamical modelling of the Pal~5 stream...
%meaning we can measure kinematic properties of the full stream...
\citep{Ibata:2017} found some RRL in pal 5, but no proper motions.
This is useful for modeling...
Many stream candidates found in RRL alone \citep{Mateu:2018}.

>>>>

RR Lyrae stars (RRLs) are pulsating horizontal branch giants intrinsically more luminous than main sequence turn-off stars. Although about $\sim 10,000$ time less numerous than main sequence stars, RRLs are extremely useful probes of the Galactic halo as they are well established as standard candles; they can be observed to large distances \citep{Medina:2018,Sesar:2017c}; and are reliably identified based on their photometric variability, with little to no contamination from other types of variable stars \citep[e.g.]{Holl:2018,Drake:2017,Mateu:2012}.  Their distances can be estimated based on photometric data alone, with errors that go from $\sim8\%$ in the optical to as low as 3\% in the infrared \citep{Neeley:2017, Sesar:2017a}.

Previous studies have found RRLs in Pal5's tails: the first two were found by \citet{Vivas:2001} in the QUEST survey; and more recently \citep{Ibata:2017} found about ten new RRLs using the catalogue of PS1 candidate RRLs from \citet{Hernitschek:2016}. At its current distance, Pal~5's horizontal branch is bright enough ($\G\sim17.3$) to be observed by \Gaia. Therefore, combining \Gaia~DR2 proper motions with the newer PS1 RRL catalogue from \citet{Sesar:2017b} --with improved classification probabilities and light curve periods--  will allows to make precise distance and proper motion measurements along the full extent of the Pal~5 stream. Both, key ingredients for dynamical modelling of the Pal~5 stream...
%meaning we can measure kinematic properties of the full stream...
\citep{Ibata:2017} found some RRL in pal 5, but no proper motions.
This is useful for modeling...
Many stream candidates found in RRL alone \citep{Mateu:2018}.
